---
title: "Signac: create data objects from cellranger quantifications, compute QC metrics"
output:
  workflowr::wflow_html:
    toc: true
editor_options:
  chunk_output_type: console
---

# Load packages and gene annotations
```{r}
library(tidyverse)
library(GenomicRanges)
library(Seurat)
library(Signac)
library(EnsDb.Hsapiens.v86)

annotations <- GetGRangesFromEnsDb(ensdb = EnsDb.Hsapiens.v86)
seqlevelsStyle(annotations) <- 'UCSC'
```


# PBMC
## Import cellranger counts and metadata
```{r echo=FALSE}
PBMC_cr_counts <- Read10X_h5(filename = "/proj/jmsimon/alevinForATAC/10X_scATAC_testData/10x_PBMC_cellranger/outs/filtered_peak_bc_matrix.h5")
```
```{r eval=FALSE}
PBMC_cr_counts <- Read10X_h5(filename = "10x_PBMC_cellranger/outs/filtered_peak_bc_matrix.h5")
```

```{r echo=FALSE}
PBMC_cr_metadata <- read.csv(
  file = "/proj/jmsimon/alevinForATAC/10X_scATAC_testData/10x_PBMC_cellranger/outs/singlecell.csv",
  header = TRUE,
  row.names = 1
)
```
```{r eval=FALSE}
PBMC_cr_metadata <- read.csv(
  file = "10x_PBMC_cellranger/outs/singlecell.csv",
  header = TRUE,
  row.names = 1
)
```

## Create chromatin assay and seurat object
```{r echo=FALSE}
PBMC_cr_chrom_assay <- CreateChromatinAssay(
  counts = PBMC_cr_counts,
  sep = c(":", "-"),
  genome = 'hg38',
  fragments = '/proj/jmsimon/alevinForATAC/10X_scATAC_testData/10x_PBMC_cellranger/outs/fragments.tsv.gz',
  min.cells = 10,
  min.features = 200
)
```

```{r eval=FALSE}
PBMC_cr_chrom_assay <- CreateChromatinAssay(
  counts = PBMC_cr_counts,
  sep = c(":", "-"),
  genome = 'hg38',
  fragments = '10x_PBMC_cellranger/outs/fragments.tsv.gz',
  min.cells = 10,
  min.features = 200
)
```
```{r}
PBMC_cr_seurat <- CreateSeuratObject(
  counts = PBMC_cr_chrom_assay,
  assay = "peaks",
  meta.data = PBMC_cr_metadata
)
```

## Add gene annotation information to the object and prepend cell type on cell barcodes
```{r}
Annotation(PBMC_cr_seurat) <- annotations

PBMC_cr_seurat$Sample <- "PBMC"
PBMC_cr_seurat <- RenameCells(PBMC_cr_seurat, add.cell.id = "PBMC")
```

# HGMM
## Import cellranger counts and metadata
```{r echo=FALSE}
HGMM_cr_counts <- Read10X_h5(filename = "/proj/jmsimon/alevinForATAC/10X_scATAC_testData/10x_HGMM_cellranger/outs/filtered_peak_bc_matrix.h5")
```

```{r eval=FALSE}
HGMM_cr_counts <- Read10X_h5(filename = "10x_HGMM_cellranger/outs/filtered_peak_bc_matrix.h5")
```

```{r echo=FALSE}
HGMM_cr_metadata <- read.csv(
  file = "/proj/jmsimon/alevinForATAC/10X_scATAC_testData/10x_HGMM_cellranger/outs/singlecell.csv",
  header = TRUE,
  row.names = 1
)
```
```{r eval=FALSE}
HGMM_cr_metadata <- read.csv(
  file = "10x_HGMM_cellranger/outs/singlecell.csv",
  header = TRUE,
  row.names = 1
)
```

## Create chromatin assay and seurat object
```{r echo=FALSE}
HGMM_cr_chrom_assay <- CreateChromatinAssay(
  counts = HGMM_cr_counts,
  sep = c(":", "-"),
  genome = 'hg38',
  fragments = '/proj/jmsimon/alevinForATAC/10X_scATAC_testData/10x_HGMM_cellranger/outs/fragments.tsv.gz',
  min.cells = 10,
  min.features = 200
)
```
```{r eval=FALSE}
HGMM_cr_chrom_assay <- CreateChromatinAssay(
  counts = HGMM_cr_counts,
  sep = c(":", "-"),
  genome = 'hg38',
  fragments = '10x_HGMM_cellranger/outs/fragments.tsv.gz',
  min.cells = 10,
  min.features = 200
)
```
```{r}
HGMM_cr_seurat <- CreateSeuratObject(
  counts = HGMM_cr_chrom_assay,
  assay = "peaks",
  meta.data = HGMM_cr_metadata
)
```

## Add gene annotation information to the object and prepend cell type on cell barcodes
```{r}
Annotation(HGMM_cr_seurat) <- annotations

HGMM_cr_seurat$Sample <- "HGMM"
HGMM_cr_seurat <- RenameCells(HGMM_cr_seurat, add.cell.id = "HGMM")
```





# Compute QC metrics
## HGMM
```{r echo=FALSE}
HGMM_cr_path <- '/proj/jmsimon/alevinForATAC/10X_scATAC_testData/10x_HGMM_cellranger/outs/fragments.tsv.gz'
```
```{r eval=FALSE}
HGMM_cr_path <- '10x_HGMM_cellranger/outs/fragments.tsv.gz'
```
```{r}
HGMM_cr_fragmentInfo <- CountFragments(HGMM_cr_path)
rownames(HGMM_cr_fragmentInfo) <- paste0(HGMM_cr_seurat$Sample,"_",HGMM_cr_fragmentInfo$CB)

# Attach cell metadata to seurat object
HGMM_cr_seurat$fragments <- HGMM_cr_fragmentInfo[colnames(HGMM_cr_seurat), "frequency_count"]
HGMM_cr_seurat$mononucleosomal <- HGMM_cr_fragmentInfo[colnames(HGMM_cr_seurat), "mononucleosomal"]
HGMM_cr_seurat$nucleosome_free <- HGMM_cr_fragmentInfo[colnames(HGMM_cr_seurat), "nucleosome_free"]
HGMM_cr_seurat$reads_count <- HGMM_cr_fragmentInfo[colnames(HGMM_cr_seurat), "reads_count"]

# Calculate FRiP
HGMM_cr_seurat <- FRiP(
  object = HGMM_cr_seurat,
  assay = 'peaks',
  total.fragments = "fragments"
)

# Calculate signal over excluded regions
HGMM_cr_seurat$blacklist_fraction <- FractionCountsInRegion(
  object = HGMM_cr_seurat, 
  assay = 'peaks',
  regions = blacklist_hg38
)

# Compute nucleosome signal score per cell
HGMM_cr_seurat <- NucleosomeSignal(HGMM_cr_seurat)

# Compute TSS enrichment
HGMM_cr_seurat <- TSSEnrichment(HGMM_cr_seurat, fast=FALSE)
```

### Plot TSS enrichment
```{r}
HGMM_cr_seurat$high.tss <- ifelse(HGMM_cr_seurat$TSS.enrichment > 2, 'High', 'Low')
TSSPlot(HGMM_cr_seurat, group.by = 'high.tss') + NoLegend()
```

### Plot nucleosome fragment length periodicity for all cells
```{r}
HGMM_cr_seurat$nucleosome_group <- ifelse(HGMM_cr_seurat$nucleosome_signal > 4, 'NS > 4', 'NS < 4')
FragmentHistogram(object = HGMM_cr_seurat, group.by = 'nucleosome_group')
```



## PBMC
```{r echo=FALSE}
PBMC_cr_path <- '/proj/jmsimon/alevinForATAC/10X_scATAC_testData/10x_PBMC_cellranger/outs/fragments.tsv.gz'
```
```{r eval=FALSE}
PBMC_cr_path <- '10x_PBMC_cellranger/outs/fragments.tsv.gz'
```

```{r}
PBMC_cr_fragmentInfo <- CountFragments(PBMC_cr_path)
rownames(PBMC_cr_fragmentInfo) <- paste0(PBMC_cr_seurat$Sample,"_",PBMC_cr_fragmentInfo$CB)

# Attach cell metadata to seurat object
PBMC_cr_seurat$fragments <- PBMC_cr_fragmentInfo[colnames(PBMC_cr_seurat), "frequency_count"]
PBMC_cr_seurat$mononucleosomal <- PBMC_cr_fragmentInfo[colnames(PBMC_cr_seurat), "mononucleosomal"]
PBMC_cr_seurat$nucleosome_free <- PBMC_cr_fragmentInfo[colnames(PBMC_cr_seurat), "nucleosome_free"]
PBMC_cr_seurat$reads_count <- PBMC_cr_fragmentInfo[colnames(PBMC_cr_seurat), "reads_count"]

# Calculate FRiP
PBMC_cr_seurat <- FRiP(
  object = PBMC_cr_seurat,
  assay = 'peaks',
  total.fragments = "fragments"
)

# Calculate signal over excluded regions
PBMC_cr_seurat$blacklist_fraction <- FractionCountsInRegion(
  object = PBMC_cr_seurat, 
  assay = 'peaks',
  regions = blacklist_hg38
)

# Compute nucleosome signal score per cell
PBMC_cr_seurat <- NucleosomeSignal(PBMC_cr_seurat)

# Compute TSS enrichment
PBMC_cr_seurat <- TSSEnrichment(PBMC_cr_seurat, fast=FALSE)
```

## Plot TSS enrichment
```{r}
PBMC_cr_seurat$high.tss <- ifelse(PBMC_cr_seurat$TSS.enrichment > 2, 'High', 'Low')
TSSPlot(PBMC_cr_seurat, group.by = 'high.tss') + NoLegend()
```

## Plot nucleosome fragment length periodicity for all cells
```{r}
PBMC_cr_seurat$nucleosome_group <- ifelse(PBMC_cr_seurat$nucleosome_signal > 4, 'NS > 4', 'NS < 4')
FragmentHistogram(object = PBMC_cr_seurat, group.by = 'nucleosome_group')
```


# Save workspace
```{r echo=FALSE}
save.image("../Cellranger_HGMM_PBMC_seurat_090222_QC.RData")
```

```{r eval=FALSE}
save.image("Cellranger_HGMM_PBMC_seurat_090222_QC.RData")
```

