---
title: "Signac: Integration and clustering of chromap quantifications"
output:
  workflowr::wflow_html:
    toc: true
editor_options:
  chunk_output_type: console
---

# Load packages and previously saved workspace
```{r}
library(tidyverse)
library(GenomicRanges)
library(Seurat)
library(Signac)
library(EnsDb.Hsapiens.v86)

```

```{r echo=FALSE}
load("../cCRE_hg38_10x_HGMM_PBMC_chromap_fragments_MACS_q01_unionPeaks_merge100_seurat_090222_QC.RData")
```
```{r eval=FALSE}
load("cCRE_hg38_10x_HGMM_PBMC_chromap_fragments_MACS_q01_unionPeaks_merge100_seurat_090222_QC.RData")
```


# Merge chromap seurat objects, subset, then split into a list
```{r}
chromap.combined <- merge(HGMM_10x_seurat,y = PBMC_10x_seurat)


# Subset for features observed in at least 20 cells
filtFeatures <- FindTopFeatures(chromap.combined,min.cutoff=20)
chromap.combined <- subset(chromap.combined, features = VariableFeatures(filtFeatures))

# Split by sample into a list
chromap.list <- SplitObject(chromap.combined, split.by="Sample")
```

# Iterate over sample list, subset based on QC metrics, and generate lsi embeddings
```{r}
for (i in 1:length(chromap.list)) {
	chromap.list[[i]] <- subset(x = chromap.list[[i]],
		features = VariableFeatures(filtFeatures), 
		subset = nCount_peaks > 1000 &
		nCount_peaks < 100000 &
		FRiP > 0.15 &
		blacklist_fraction < 0.05 &
		nucleosome_signal < 4 &
		TSS.enrichment > 2
		)

	chromap.list[[i]] <- FindTopFeatures(chromap.list[[i]], min.cutoff = 'q0')
	chromap.list[[i]] <- RunTFIDF(chromap.list[[i]], assay="peaks")
	chromap.list[[i]] <- RunSVD(chromap.list[[i]])
	chromap.list[[i]] <- RunUMAP(chromap.list[[i]], reduction = "lsi", dims = 2:30)
}


# Subset and process the combined dataset
chromap.combined <- subset(
  x = chromap.combined,
  features = VariableFeatures(filtFeatures), 
  subset = nCount_peaks > 1000 &
    nCount_peaks < 100000 &
    FRiP > 0.15 &
    blacklist_fraction < 0.05 &
    nucleosome_signal < 4 &
    TSS.enrichment > 2
)
```

# Get filtered dimensions
```{r}
table(chromap.combined$Sample)

lapply(chromap.list, dim)

```

# Repeat here on the combined object
```{r}
chromap.combined <- FindTopFeatures(chromap.combined, min.cutoff = 'q0')
chromap.combined <- RunTFIDF(chromap.combined, assay="peaks")
chromap.combined <- RunSVD(chromap.combined)
chromap.combined <- RunUMAP(chromap.combined, reduction = "lsi", dims = 2:30)
```

# Integrate all data
```{r}
# Find integration anchors
integration.anchors <- FindIntegrationAnchors(object.list = chromap.list, 
	anchor.features = rownames(chromap.combined), 
	reduction = "rlsi", 
	dims = 2:30
)

# Integrate LSI embeddings
# Arbitrarily reduced k.weight to 50 to get around "number of items to replace is not a multiple of replacement length" error
chromap.integrated <- IntegrateEmbeddings(anchorset = integration.anchors, 
	reductions = chromap.combined[["lsi"]],
	new.reduction.name = "integrated_lsi",
	dims.to.integrate = 1:30,
	k.weight = 50
)
```

# Create a new UMAP using the integrated embeddings
```{r}
chromap.integrated <- RunUMAP(chromap.integrated, reduction = "integrated_lsi", dims = 2:30)
DimPlot(chromap.integrated, group.by = "Sample")
```

# Identify clusters
```{r}
chromap.integrated <- FindNeighbors(object = chromap.integrated, reduction = 'integrated_lsi', dims = 2:30)
chromap.integrated <- FindClusters(object = chromap.integrated, verbose = FALSE, algorithm = 3, resolution = 0.5)
DimPlot(object = chromap.integrated, label = TRUE) + NoLegend()
```


# Compute gene activity matrix and insert this as a pseudo RNA assay
```{r}
gene.activities <- GeneActivity(chromap.integrated)

# Add the gene activity matrix to the Seurat object as a new assay and normalize it
chromap.integrated[['RNA']] <- CreateAssayObject(counts = gene.activities)
chromap.integrated <- NormalizeData(
  object = chromap.integrated,
  assay = 'RNA',
  normalization.method = 'LogNormalize',
  scale.factor = median(chromap.integrated$nCount_RNA)
)

DefaultAssay(chromap.integrated) <- 'RNA'
```

# Plot the same marker genes from Seurat vignette
```{r}
FeaturePlot(
  object = chromap.integrated,
  features = c('MS4A1', 'CD3D', 'LEF1', 'NKG7', 'TREM1', 'LYZ'),
  pt.size = 0.1,
  max.cutoff = 'q95',
  ncol = 3
)
```

## Save workspace
```{r echo=FALSE}
save.image("../cCRE_hg38_10x_HGMM_PBMC_chromap_fragments_MACS_q01_unionPeaks_merge100_seurat_090222_QC_integrated.RData")
```

```{r eval=FALSE}
save.image("cCRE_hg38_10x_HGMM_PBMC_chromap_fragments_MACS_q01_unionPeaks_merge100_seurat_090222_QC_integrated.RData")
```

